\bf A Hierarchical Model for Heterogenous Reliability Field Data

### 2018-04-10 paper/m-lb-m-technometrics.tex
Paper and bibtek edits after 4/8 meeting

\noindent%
{\it Keywords:} GLFP, Bathtub hazard, Censored data, Hierarchical models, Bayesian estimation
\vfill

@@

\subsection{Weibull distribution parameterization}
\label{sec:Weibull parameterization}
The Weibull cumulative
 distribution function (cdf) is
\begin{align}
\label{equation:weibull.cdf}
\Pr(T \leq t|\alpha,\beta ) &= F(t|\alpha,\beta)=1-
\exp \left [-\left (\frac{t}{\alpha} \right )^{\beta}
\right ], \quad t > 0,
\end{align}
where $\beta>0$ is the Weibull shape parameter and $\alpha>0$ is a
scale parameter. Because $\log(T)$ has a smallest extreme value
distribution (a member of the location-scale family of
distributions), the Weibull cdf can also be written as
\begin{align*}
\Pr(T \leq t| \mu,\sigma ) &= F(t| \mu,\sigma)= \Phi_{SEV}\left[\frac{\log(t)-\mu}{\sigma}\right], \quad t > 0,
\end{align*}
where $\Phi_{SEV}(z)=1-\exp[-\exp(z)]$ is the standard smallest extreme value
distribution cdf and
$\mu=\log(\alpha)$ and $\sigma=1/\beta$ are, respectively, location
and scale parameters for the distribution of $\log(T)$. Therefore, the Weibull
distribution is a member of the log-location-scale family of
distributions.

We will use an alternative parameterization
where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$. Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are two important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}

@@

\subsection{Model selection}
From the point of view of a consumer or purchaser, the goal of an analysis such as this is may be to rate manufacturers and/or to inform future purchases based on predicted product performance. Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.  Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
posterior draws from a fitted model.  The log pointwise predictive density (lpd), is one method to quantify predictive accuracy,

\begin{align*}
\mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
\end{align*}
\noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$. Notationally suppressing the conditioning on $t^L$ and $c$,
\begin{equation}
\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
\label{elpd}
\end{equation}
where $h$ is a density for the true distribution of $t_{new,g,i}$. Although $h$ is unknown, leave-one-out cross-validation provides a robust estimator:
\begin{equation}
\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
\label{loo-est}
\end{equation}
The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent. It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $p(t_{g,i}|t_{-(g,i)})$ with $\sum_{s=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$. In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result. For details, refer to \cite{vehtari}.
%When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
% To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.  As outlined in Vehtari et. al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.  For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.  
% The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).

>>>>

\noindent%
{\it Keywords:} GLFP, Bathtub hazard, Censored data, Bayesian estimation, Stan
\vfill

@@

\subsection{Weibull distribution parameterization}
\label{sec:Weibull parameterization}
The Weibull cumulative
 distribution function (cdf) is
\begin{align}
\label{equation:weibull.cdf}
\Pr(T \leq t|\alpha,\beta ) &= F(t|\alpha,\beta)=1-
\exp \left [-\left (\frac{t}{\alpha} \right )^{\beta}
\right ], \quad t > 0,
\end{align}
where $\beta>0$ is the Weibull shape parameter and $\alpha>0$ is a
scale parameter. Because $\log(T)$ has a smallest extreme value
distribution (a member of the location-scale family of
distributions), the Weibull cdf can also be written as
\begin{align*}
\Pr(T \leq t| \mu,\sigma ) &= F(t| \mu,\sigma)= \Phi_{SEV}\left[\frac{\log(t)-\mu}{\sigma}\right], \quad t > 0,
\end{align*}
where $\Phi_{SEV}(z)=1-\exp[-\exp(z)]$ is the standard smallest extreme value
distribution cdf and
$\mu=\log(\alpha)$ and $\sigma=1/\beta$ are, respectively, location
and scale parameters for the distribution of $\log(T)$. In addition to showing the Weibull
distribution is a member of the log-location-scale family of
distributions, this step highlights the generalizability of our approach to other distributions 
that fall into this class.

Going forward we use an alternative parameterization
where $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$. Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are two important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}

@@

\subsection{Model selection}
Bayes factors is a classical apporach to compare alternative model specifications.  However, they can be problematic, especially in high dimensional settings where there is no concensus on how to properly compute the required marginal likelihood \citep{robert2016}. Moreover, from the point of view of a consumer or purchaser, the goal of an analysis such as this is to rate manufacturers and/or to inform future purchases based on predicted product performance. Therefore, we should compare model specifications based on accurate prediction for future units of previously observed products.  Leave-one-out cross-validation (LOO) is a method to estimate out-of-sample predictive accuracy using
posterior draws from a fitted model.  The log pointwise predictive density (lpd), is one method to quantify predictive accuracy,

\begin{align*}
\mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
\end{align*}
\noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$. Notationally suppressing the conditioning on $t^L$ and $c$,
\begin{equation}
\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
\label{elpd}
\end{equation}
where $h$ is a density for the true distribution of $t_{new,g,i}$. Although $h$ is unknown, leave-one-out cross-validation provides a robust estimator:
\begin{equation}
\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
\label{loo-est}
\end{equation}
The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent. It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $p(t_{g,i}|t_{-(g,i)})$ with $\sum_{s=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$. In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.  A measure of the effective number of parameters is also available using the sum of the posterior variance of the log predictive density for each observation.  For details, refer to \cite{vehtari}.
%When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
% To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.  As outlined in Vehtari et. al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.  For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.  
% The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).
